---
title: "Exam"
author: "Sau-Chin Chen"
date: "2016年10月31日"
output: html_document
---
1. An equivalence test using the TOST procedure indicates the observed effect is significantly smaller than the upper equivalence bound, and significantly larger than the lower equivalence bound. Traditional NHST also shows the effect is statistically significant. Which conclusion can you draw?

A) The effect is significant, and large enough to be worthwhile to examine.

**B)** The effect is significant, but too small to be worthwhile to examine.

C) The effect is undetermined; we can not conclude it is significant, nor too small to be worthwhile to examine.

D) The effect is too small to be worthwhile to examine, and not reliably different from 0.

2. Open the TOST t-test spreadsheet. Researchers examined how much fun students found lectures when they only saw the teacher in videos, compared to how much fun the lectures were when they saw the teacher in real life. They found doing the course was about as much fun when the students saw the lecturer in real life (M1 = 4.88, SD1 = 1.35, n1 = 180) as when they saw the lecturer in the videos (M2 = 5.01, SD2 = 1.21, n2 = 190). The researchers considered an effect size of Cohen’s d = 0.2 too small to be worthwhile. What can the researchers conclude?


A) The effect was too small to be worthwhile.

B) The effect was significantly different from 0.

**C)** The effect was undetermined: not significant and not equivalent.

3. A researcher wants to design a study where she has enough power to determine there is no worthwhile effect. She performs a power analysis for an equivalence test for an independent t-test. She desired 95% power, uses an alpha of 0.10, and sets the lower equivalence bound to d = -0.25 and the upper equivalence bound to 0.25. Assuming the true effect is 0, how many participants would she need in total?

A) 337

B) 416

**C)** 674

D) 832

4. Open the Bayes-Factor One-Sided t-test.R script. A researcher named Chris compared whether students performed better on a quiz when they had to perform this quiz in the morning, or in the afternoon. The class consisted of 250 students, and the observed difference was Cohen’s dz = 0.05. The researcher had a-priori expected a reasonable effect of time of day on grades, and had a prior for Cohen’s dz of 0.3, and a prior on the sd of 0.2. Which conclusion can the researcher draw?
```{r Exam4}
#This code is made by Jeff Rouder: http://jeffrouder.blogspot.nl/2016/01/what-priors-should-i-use-part-i.html
#This code is for a one-sided t-test, testing a difference against 0.

N <- 250 #Enter sample size
dz <- 0.05 #Enter Cohen's dz effect size observed in the dependent t-test or one-sided t-test

dz_prior <- 0.3 #Enter effect size dz for the prior 
sd_prior <- 0.2 #Enter sd of the effect sizes of the prior - the higher, the wider the prior is

lo <- -Inf #lower bound of support (e.g., set to 0 if effects < 0 is not possible)
hi <- Inf #upper bound of support

#specify prior
altDens=function(delta) 
  dnorm(delta,dz_prior,sd_prior)*as.integer(delta>lo)*as.integer(delta<hi)
#Normalize alternative density in case user does not, 
K=1/integrate(altDens,lower=lo,upper=hi)$value
f=function(delta) K*altDens(delta)

delta=seq(-3,3,.01)
#Plot Alternative as a density and Null as a point arrow
#png(file=paste('prior.png'),width=6000,height=4000, res = 1000)
maxAlt=max(f(delta))
plot(delta,f(delta),typ='n',xlab="Effect Size Parameter Delta",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
lines(delta,f(delta),col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
#dev.off()

#Prediction Function Under Null
nullPredF=function(obs,N) dt(sqrt(N)*obs,N-1)

#Prediction Function Under the Alternative
altPredIntegrand=function(delta,obs,N) 
  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*f(delta)
altPredF=function(obs,N) 
  integrate(altPredIntegrand,lower=lo,upper=hi,obs=obs,N=N)$value

obs=delta
I=length(obs)
nullPred=nullPredF(obs,N)
altPred=1:I
for (i in 1:I) altPred[i]=altPredF(obs[i],N)

#Evaluate Predicted Density at Observed Value dz
valNull=nullPredF(dz,N)
valAlt=altPredF(dz,N)

#Plot The Predictions
#png(file=paste('posterior.png'),width=6000,height=4000, res = 1000)
top=max(altPred,nullPred)
plot(type='l',obs,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main=paste("Bayes factor (alt/null) is ",round(valAlt/valNull,digits =3)),col='darkblue',lwd=2)
lines(obs,altPred,col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
abline(v=dz,lty=2,lwd=2,col='red')
points(pch=19,c(dz,dz),c(valNull,valAlt))
#dev.off()

cat("Bayes factor (alt/null) is ",valAlt/valNull,", the t-value is ",sqrt(N)*dz," and the p-value is",2*(1-pt(abs(sqrt(N)*dz),N-1)))

#? Daniel Lakens, 2016. 
# This work is licensed under a Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License. https://creativecommons.org/licenses/by-nc-sa/4.0/
```

A) The alternative hypothesis has become more likely, given the data, with a $BF_10$ = 0.203.

**B)** The null hypothesis has become more likely, given the data, with a $BF_10$ = 0.203.

C) The alternative hypothesis has become more likely, given the data, with a BF10 = 0.25.

D) The null hypothesis has become more likely, given the data, with a BF10 = 0.25.

5. Lisa, a colleague of researcher Chris in the previous question had a-priori predicted there would be a much smaller effect of time of day at all, having a prior for Cohen’s dz of 0.1. Given a prior on the sd of 0.2, and the observed data (a Cohen’s dz of 0.05 in a sample size of 250 students) what would Lisa conclude?
```{r Exam5}
#This code is made by Jeff Rouder: http://jeffrouder.blogspot.nl/2016/01/what-priors-should-i-use-part-i.html
#This code is for a one-sided t-test, testing a difference against 0.

N <- 250 #Enter sample size
dz <- 0.05 #Enter Cohen's dz effect size observed in the dependent t-test or one-sided t-test

dz_prior <- 0.1 #Enter effect size dz for the prior 
sd_prior <- 0.2 #Enter sd of the effect sizes of the prior - the higher, the wider the prior is

lo <- -Inf #lower bound of support (e.g., set to 0 if effects < 0 is not possible)
hi <- Inf #upper bound of support

#specify prior
altDens=function(delta) 
  dnorm(delta,dz_prior,sd_prior)*as.integer(delta>lo)*as.integer(delta<hi)
#Normalize alternative density in case user does not, 
K=1/integrate(altDens,lower=lo,upper=hi)$value
f=function(delta) K*altDens(delta)

delta=seq(-3,3,.01)
#Plot Alternative as a density and Null as a point arrow
#png(file=paste('prior.png'),width=6000,height=4000, res = 1000)
maxAlt=max(f(delta))
plot(delta,f(delta),typ='n',xlab="Effect Size Parameter Delta",ylab="Density",ylim=c(0,1.4*maxAlt),main="Models")
arrows(0,0,0,1.3*maxAlt,col='darkblue',lwd=2)
lines(delta,f(delta),col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
#dev.off()

#Prediction Function Under Null
nullPredF=function(obs,N) dt(sqrt(N)*obs,N-1)

#Prediction Function Under the Alternative
altPredIntegrand=function(delta,obs,N) 
  dt(sqrt(N)*obs,N-1,sqrt(N)*delta)*f(delta)
altPredF=function(obs,N) 
  integrate(altPredIntegrand,lower=lo,upper=hi,obs=obs,N=N)$value

obs=delta
I=length(obs)
nullPred=nullPredF(obs,N)
altPred=1:I
for (i in 1:I) altPred[i]=altPredF(obs[i],N)

#Evaluate Predicted Density at Observed Value dz
valNull=nullPredF(dz,N)
valAlt=altPredF(dz,N)

#Plot The Predictions
#png(file=paste('posterior.png'),width=6000,height=4000, res = 1000)
top=max(altPred,nullPred)
plot(type='l',obs,nullPred,ylim=c(0,top),xlab="Observed Effect Size",ylab="Density",main=paste("Bayes factor (alt/null) is ",round(valAlt/valNull,digits =3)),col='darkblue',lwd=2)
lines(obs,altPred,col='darkgreen',lwd=2)
legend("topright",legend=c("Null","Alternative"),col=c('darkblue','darkgreen'),lwd=2)
abline(v=dz,lty=2,lwd=2,col='red')
points(pch=19,c(dz,dz),c(valNull,valAlt))
#dev.off()

cat("Bayes factor (alt/null) is ",valAlt/valNull,", the t-value is ",sqrt(N)*dz," and the p-value is",2*(1-pt(abs(sqrt(N)*dz),N-1)))

#? Daniel Lakens, 2016. 
# This work is licensed under a Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International License. https://creativecommons.org/licenses/by-nc-sa/4.0/
```

A) The alternative hypothesis has become more likely, but Lisa’s belief in the alternative hypothesis increases less than Chris’ belief.

**B)** The null hypothesis has become more likely, but Lisa’s belief in the null hypothesis increases less than Chris’ belief.

C) The alternative hypothesis has become more likely, and Lisa’s belief in the alternative hypothesis increases more than Chris’ belief.

D) The null hypothesis has become more likely, but Lisa’s belief in the null hypothesis increases more than Chris’ belief.
